Multi-Agent Path Finding with Matching using Increasing Cost Tree Search

### 2021-06-16 docs/paper.tex
Added references to other algorithms used as starting point

	\begin{wrapfigure}[19]{R}{.25\textwidth}
		\begin{minipage}{\linewidth}
			\centering\captionsetup[subfigure]{justification=centering}
			\includegraphics[width=\linewidth]{img/vertex-conflict}
			\subcaption{Vertex conflict}
			\label{fig:conflictsa}\par\vfill
			\includegraphics[width=0.75\linewidth]{img/edge-conflict}
			\subcaption{Edge conflict}
			\label{fig:conflictsb}
		\end{minipage}
		\vspace{-10pt}
		\caption{The two types of conflict (from Fig. 1 in \cite{stern2019})}\label{fig:conflicts}
	\end{wrapfigure}
	This work is the outcome of a search for algorithms for optimally solving MAPFM based on ICTS as an alternative to the approach outlined above\footnote{To the knowledge of the author, to date, no such algorithm is described in the literature}. Two classes of algorithms were found, one based on exhaustive enumeration of matchings using a variant of ICTS as subroutine, the other based on modifications to the ICT structure and its search. In the context of a research project in which multiple MAPF algorithms in addition to ICTS were taken as starting points for MAPFM algorithms, one exhaustive ICTS algorithm is compared to these alternatives on a set of benchmarks. In addition, properties of the various algorithm variants are discussed and their performance is compared experimentally.

@@

	In implementation, ordered enumeration is approximated using a priority queue that maintains a subset of matchings of a fixed size $M_{pq}$\footnote{$M_{pq} = 10000$ was used in implementation but this could be much increased at the cost of memory}. Algorithm \ref{pq-algo} outlines the procedure. This way of getting around memory limitations originates in [TODO: cite Jaap's paper].

@@

	\restoregeometry
	\subsection{Experimental setup}
	Because MAPFM lies on the fringe of pathfinding research, no standard benchmark exists. For MAPF, the Moving AI benchmarks \cite{sturtevant2012} are often used and potentially a new MAPFM standard benchmark could be defined by generating suitable team-partitions of agents for these MAPF instances; however, it was noticed that optimally solving even small MAPFM instances with comparatively few agents and short distances to goals was computationally already a challenge. This observation informed the creation of a map generator within the peer group with various parameters. Using this map generator, two types of 20x20 grids were generated:
	\begin{enumerate}
		\item 25: a grid with 25\% of tiles being obstacles.
		\item 75: a grid with 75\% of tiles being obstacles. 
	\end{enumerate}
	For each of these types, two sets of maps were generated with between $1$ to $16$ agents and 200 instances generated for each such agent-setting: one set where all agents belong to one team, another in which the agents were divided into three teams (or $1,2$ teams strictly speaking for the datapoints with respectively $1,2$ agents).

	The first experiments were performed on an Intel i7-8750H hexacore operating at 2.2GHz with 16GiB of memory. Here, for each setting, a fixed subset of 50 out of 200 maps were used.

	The second experiments comparing the different MAPFM algorithms within the peer group were run on a server with a Intel Xeon E5-2683 dodecacore operating at 2.0Ghz with 8GiB of RAM. All 200 maps were used for each setting. Although these algorithms have different authors, all are implemented in Python 3 (with one using an ILP solver implemented in C++ as a subroutine) to avoid performance differences as often seen between interpreted versus compiled languages. Nonetheless, conclusions are here a bit more limited, as performance differences are sometimes subtle and could potentially be explained by slight implementation optimizations rather than algorithmic properties.

	\subsection{Experiment 1: ICTS-based MAPFM algorithms}
	Figure \ref{fig:i-probs} reports the fraction of instances that each algorithm variant solved within the timeout of 120s in different settings. The meaning of the abbreviations used in the legend are as follows:
	\begin{itemize}
		\begin{minipage}{0.45\linewidth}
			\item S: Simple pruning
			\item E: Enhanced pruning
			\item C: Child pruning
		\end{minipage}
		\begin{minipage}{0.45\linewidth}
			\item ID: Independence detection
			\item B: Bounded search
			\item O: Approximate ordered enumeration
		\end{minipage}
	\end{itemize}
	\subsubsection{Analysis}
	Across scenarios, it is seen that typically exhaustive methods outperform ICTS-m algorithms (\ref{fig:i-probs}), with performance being more similar in the 75 maps with more obstructions. In particular on 75 maps with 3 teams (\ref{fig:i-75-3-p}), the performance of ICTS-m+ID (and optional pruning) is seen to outperform bounded exhaustive search as $k$ increases, and some of these also manage to solve a small fraction of problems of 75 maps with 1 team that exhaustive methods cannot within the time-out (\ref{fig:i-75-1-p}).

	Between exhaustive variants, with exception of the ID variant, a type of hierarchy can be distinguished, best illustrated by \ref{fig:i-25-3-p}; almost without exception, Exh+S+C+B+O outperforms Exh+E+B+O which outperforms Exh+E+B which in turn outperforms Exh+E. Bounded search outperforms unbounded search as expected, given that the former does at most as much work as the latter and often less. Approximate ordering outperforming unordered enumeration is not quite as obvious from a theoretical standpoint, as discussed in \ref{ordered-enum}. Simple pruning with child pruning outperforming exhaustive pruning shows the potential of this novel method. In particular as $k$ grows larger and in scenarios with more obstacles (\ref{fig:i-75-1-p},\ref{fig:i-75-3-p}), it appears that having to search a smaller ICT is more important in terms of performance than being able to skip the $k$-agent low-level search more often. In the

	The benefit of using ID together with ordered enumeration is less clear. Ordered enumeration with ID performs much better in \ref{fig:i-75-3-p}, a setting with more obstacles and more teams, but ordered enumeration without ID on the other hand is superior in an open setting with a single team (\ref{fig:i-25-1-p}). The overhead associated with the various planning and merging steps (see Algorithm \ref{id-algo}) could be the cause the slowdown observed once $M$ grows large (\ref{fig:i-25-1-p},\ref{fig:i-75-1-p}); once the upper bound is tightened, ICTs only have to be searched shallowly so there is no point in going through the steps of ID. This explains why no speedup is seen, as ICTS-m between and ICTS-m+ID.
	\subsection{Experiment 2: Algorithms for MAPFM}
	Figure \ref{fig:r-probs} reports the fraction of instances solved within 120s by various algorithms for MAPFM. The algorithms are as follows
	\begin{itemize}
		\begin{minipage}{0.49\linewidth}
			\item A*+ID+OD with matching ID and sorted exhaustive matching\cite{bruin2021}
			\item EPEA* with matching ID and sorted exhaustive matching\cite{jong2021}
			\item M* with sorted exhaustive matching\cite{donszelmann2021}
		\end{minipage}
		\hspace{1cm}
		\begin{minipage}{0.40\linewidth}
			\item CBM adapted to use SIC, using the successive shortest path algorithm (SSP) \cite{goldberg1987} as subroutine for solving teams\cite{baauw2021}.
			\item ICTS with approximate sorted exhaustive matching (Exh+E+B+O)

		\end{minipage}
	\end{itemize}
	One of the first things that meets the eye is the staggering performance of CBM in all scenarios with one team (\ref{fig:r-25-1-p},\ref{fig:r-75-1-p}), which are equivalent to anonymous MAPF problems. This can adequately be explained by CBM solving single teams as anonymous MAPF instances in polynomial time, meaning that solving a single team reduces to a polynomial-time procedure (SSP to be specific). It is expected that conform the results reported in \cite{ma2016}, as the number of teams approaches $k$, CBM performance will deteriorate and thus become more similar to the results for exhaustive methods here reported.
	\section{Discussion}
	Although no bound-value data was recorded, one explanation might be that often, early on in the search, single matchings with low $\Delta_m$ are found, causing the rest of the searches to be significantly curtailed. Mo
	However, closer analysis of performance differences between these pruning strategies and alternatives  (repeated enhanced pruning or pruning with pairs, quadruples, etc.) across different scenarios is something for further research.
%	Still, some implementations might be more optimized than others so this might be a factor in performance characteristics and in particular in relative performance differences.
	\section{Conclusions}
	The goal of this research was to seek alternative approaches to solving MAPFM based on ICTS. In this, the research has succeeded, but the scalability of the algorithms leaves something to be desired. Exhaustive methods, like those discussed in detail in \ref{exhaustive}, cannot be expected to scale well as teams grow larger due to their complexity, no matter what optimizations are put in place. Nonetheless, as was shown in Section \ref{experiments}, the more advanced exhaustive ICTS algorithms consistently outperform the ICTS-m based algorithms. It was also seen that in turn a SIC-adaptation of the CBM method \cite{ma2016} greatly outperforms exhaustive alternatives in the settings that were tried out. Therefore, it can be concluded that the method that arguably has the most potential to scale well, ICTS-m, has a long way to go before being usable in practical scenarios with significant problem sizes.
	\section{Future work}
	In this project, there were many roads not taken. In particular for the ICTS-m variant, there are still many things that could be experimented with in order to improve performance, such as more extensive or multi-staged pruning procedures or ways to exploit the fact that agents are partitioned into teams. For instance, the hungarian algorithm could be used to find lower-bounds for per-team SICs, which allows skipping the low-level search in many cases and can trivially be combined with pruned child generation branching within teams.

@@

	Another subject for future research might be a closer theoretical and experimental analysis of simple child pruning as alternative to enhanced pruning in the context of ICTS for MAPF, assuming this method has not been described before. Although experimentally it was shown to outperform enhanced pruning in exhaustive matching, ideally the two pruning strategies should be tested using the Moving AI benchmark \cite{sturtevant2012} to reach definite conclusions about how they compare. Using simple pruning to prune children and enhanced pruning for pruning low-level search could also be interesting to explore, as it could potentially combine the benefits of searching a smaller ICT and performing fewer $k$-agent low-level searches at the cost of spending more time pruning.
	\section{Responsible Research}
	The implementation of the ICTS-based algorithms described in this work is publicly available online in the form of a Github repository \footnote{\url{https://github.com/tbvanderwoude/icts-m}}. Included is also code for benchmarking different configurations of the ICTS solver and the maps that were used in the benchmarks reported above.

>>>>

	\begin{wrapfigure}[19]{R}{.25\textwidth}
		\begin{minipage}{\linewidth}
			\centering\captionsetup[subfigure]{justification=centering}
			\includegraphics[width=\linewidth]{img/vertex-conflict}
			\subcaption{Vertex conflict}
			\label{fig:conflictsa}\par\vfill
			\includegraphics[width=0.75\linewidth]{img/edge-conflict}
			\subcaption{Edge conflict}
			\label{fig:conflictsb}
		\end{minipage}
		\vspace{-10pt}
		\caption{The two types of conflict (from Fig. 1 in \cite{stern2019})}\label{fig:conflicts}
	\end{wrapfigure}
	This work is the outcome of a search for algorithms for optimally solving MAPFM based on ICTS as an alternative to the approach outlined above\footnote{To the knowledge of the author, to date, no such algorithm is described in the literature}, which took place as part of a project in which  other algorithms addition to ICTS were taken as starting points for MAPFM algorithms. Two classes of ICTS-based algorithms were found: one based on exhaustive enumeration of matchings using a variant of ICTS as subroutine, the other based on modifications to the ICT structure and its search. One exhaustive ICTS algorithm is compared to algorithms based on M*\cite{wagner2011}, A*+OD+ID\cite{standley2010}, CBM\cite{ma2016} and EPEA*\cite{goldenberg2014} respectively to put ICTS methods for MAPFM into a broader context.

@@

	In implementation, ordered enumeration is approximated using a priority queue that maintains a subset of matchings of a fixed size $M_{pq}$\footnote{$M_{pq} = 10000$ was used in implementation but this could be much increased at the cost of memory}. Algorithm \ref{pq-algo} outlines the procedure.

@@

	\restoregeometry
	\subsection{Experimental setup}
	Because MAPFM lies on the fringe of pathfinding research, no standard benchmark exists. For MAPF, the Moving AI benchmarks \cite{sturtevant2012} are often used and potentially a new MAPFM standard benchmark could be defined by generating suitable team-partitions of agents for these MAPF instances; however, it was noticed that optimally solving even small MAPFM instances with comparatively few agents and short distances to goals was computationally already a challenge. This observation informed the creation of a map generator by the author of \cite{jong2021} with various parameters. Using this map generator, two types of 20x20 grids were generated:
	\begin{enumerate}
		\item 25: a grid with 25\% of tiles being obstacles.
		\item 75: a grid with 75\% of tiles being obstacles. 
	\end{enumerate}
	For each of these types, two sets of maps were generated with between $1$ to $25$ agents and 200 instances generated for each such agent-setting: one set with one team, another with three teams of approximately the same size ($1,2$ teams strictly speaking for the datapoints with $1,2$ agents). For each scenario, the number of agents was increased until 0\% of instances was solved at which point no more data was recorded for greater numbers of agents.

	The first experiments were performed on an Intel i7-8750H hexacore operating at 2.2GHz with 16GiB of memory. For each setting, a fixed subset of 50 out of 200 maps were used.

	The second experiments comparing the different MAPFM algorithms within the peer group were run on a server with a virtualized Intel Xeon E5-2683 dodecacore operating at 2.0Ghz with 8GiB of RAM. All 200 maps were used for each setting.

	Although these algorithms have different authors, all are implemented in Python 3 (with one using an ILP solver implemented in C++ as a subroutine) to avoid performance differences as often seen between interpreted versus compiled languages. Nonetheless, conclusions are here a bit more limited, as performance differences are sometimes subtle and could potentially be explained by slight implementation optimizations rather than algorithmic properties.

	\subsection{Experiment 1: ICTS-based MAPFM algorithms}
	Figure \ref{fig:i-probs} reports the fraction of instances that each algorithm variant solved within the timeout of 120s in different settings. The meaning of the abbreviations used in the legend are as follows:
	\begin{itemize}
		\begin{minipage}{0.45\linewidth}
			\item S: Simple pruning
			\item E: Enhanced pruning
			\item C: Child pruning
		\end{minipage}
		\begin{minipage}{0.45\linewidth}
			\item ID: Independence detection
			\item B: Bounded search
			\item O: Approximate ordered enumeration
		\end{minipage}
	\end{itemize}
	Across scenarios, it is seen that typically exhaustive methods outperform ICTS-m algorithms (\ref{fig:i-probs}), with performance being more similar in the 75 maps with more obstructions. In particular on 75 maps with 3 teams (\ref{fig:i-75-3-p}), the performance of ICTS-m+ID (and optional pruning) is seen to outperform bounded exhaustive search as $k$ increases, and some of these also manage to solve a small fraction of problems of 75 maps with 1 team that exhaustive methods cannot within the time-out (\ref{fig:i-75-1-p}).

	Between exhaustive variants, with exception of the ID variant, a type of hierarchy can be distinguished, best illustrated by \ref{fig:i-25-3-p}; almost without exception, Exh+S+C+B+O outperforms Exh+E+B+O which outperforms Exh+E+B which in turn outperforms Exh+E. Bounded search outperforms unbounded search as expected, given that the former does at most as much work as the latter and often less. Approximate ordering outperforming unordered enumeration is not quite as obvious from a theoretical standpoint, as discussed in \ref{ordered-enum}. Simple pruning with child pruning outperforming exhaustive pruning shows the potential of this novel method. In particular as $k$ grows larger and in scenarios with more obstacles (\ref{fig:i-75-1-p},\ref{fig:i-75-3-p}), it appears that having to search a smaller ICT is more important in terms of performance than being able to skip the $k$-agent low-level search more often.

	The benefit of using ID together with ordered enumeration is less clear. Ordered enumeration with ID performs much better in \ref{fig:i-75-3-p}, a setting with more obstacles and more teams, but ordered enumeration without ID on the other hand is superior in an open setting with a single team (\ref{fig:i-25-1-p}). The overhead associated with the various planning and merging steps (see Algorithm \ref{id-algo}) could be the cause the slowdown observed once $M$ grows large (\ref{fig:i-25-1-p},\ref{fig:i-75-1-p}); once the upper bound is tightened, ICTs only have to be searched shallowly so there is no point in going through the steps of ID. This explains why no speedup is seen, as ICTS-m between and ICTS-m+ID.
	\subsection{Experiment 2: Algorithms for MAPFM}
	Figure \ref{fig:r-probs} reports the fraction of instances solved within 120s by various algorithms for MAPFM. The algorithms are as follows
	\begin{itemize}
		\begin{minipage}{0.49\linewidth}
			\item A*+ID+OD with matching ID and sorted exhaustive matching\cite{bruin2021}
			\item EPEA* with matching ID and sorted exhaustive matching\cite{jong2021}
			\item M* with sorted exhaustive matching\cite{donszelmann2021}
		\end{minipage}
		\hspace{1cm}
		\begin{minipage}{0.40\linewidth}
			\item CBM adapted to use SIC, using the successive shortest path algorithm (SSP) \cite{goldberg1987} as subroutine for solving teams\cite{baauw2021}.
			\item ICTS with approximate sorted exhaustive matching (Exh+E+B+O)
		\end{minipage}
	\end{itemize}
	With exception of CBM, all methods described use a variant of sorted exhaustive matching. This is due to the fact that from each family of algorithms, one of the best performing algorithms was chosen, which always was such an exhaustive method as opposed to a generalization of the base algorithm to MAPFM (ICTS-m in this work). A*+ID+OD and EPEA* also make use of matching ID, which is ID applied to the MAPFM instance using the respective exhaustive algorithm as MAPFM solver.

	One of the first things that meets the eye is the staggering performance of CBM in all scenarios with one team (\ref{fig:r-25-1-p},\ref{fig:r-75-1-p}), which are equivalent to anonymous MAPF problems. This can adequately be explained by CBM solving single teams as anonymous MAPF instances in polynomial time, meaning that solving a single team reduces to a polynomial-time procedure (SSP to be specific). It is expected that conform the results reported in \cite{ma2016}, as the number of teams approaches $k$, CBM performance will deteriorate and thus become more similar to the results for exhaustive methods here reported.
	\section{Discussion}
	Although no bound-value data was recorded, one explanation might be that often, early on in the search, single matchings with low $\Delta_m$ are found, causing the rest of the searches to be significantly curtailed. Mo
	However, closer analysis of performance differences between these pruning strategies and alternatives  (repeated enhanced pruning or pruning with pairs, quadruples, etc.) across different scenarios is something for further research.
%	Still, some implementations might be more optimized than others so this might be a factor in performance characteristics and in particular in relative performance differences.
	\section{Conclusions}
	The goal of this research was to seek alternative approaches to solving MAPFM based on ICTS. In this, the research has succeeded, but the scalability of the algorithms leaves something to be desired. Exhaustive methods, like those discussed in detail in \ref{exhaustive}, cannot be expected to scale well as teams grow larger due to their complexity, no matter what optimizations are put in place. Nonetheless, as was shown in Section \ref{experiments}, the more advanced exhaustive ICTS algorithms consistently outperform the ICTS-m based algorithms. Compared to the SIC-adaptation of the CBM method \cite{baauw2021} (which is adapted from \cite{ma2016}, the state-of-the-art MAPFM with makespan algorithm), it was seen that exhaustive methods in general perform quite poorly if there are fewer teams and fewer obstacles, while being competitive once teams grow smaller and maps become more constrained.

	It was also seen that in turn a SIC-adaptation of the CBM method \cite{ma2016} greatly outperforms exhaustive alternatives in the settings that were tried out. Therefore, it can be concluded that the method that arguably has the most potential to scale well, ICTS-m, has a long way to go before being usable in practical scenarios with significant problem sizes.
	\section{Future work}
	In this project, there were many roads not taken. In particular for the ICTS-m variant, there are still many things that could be experimented with in order to improve performance, such as more extensive or multi-staged pruning procedures or ways to exploit the fact that agents are partitioned into teams. For instance, the hungarian algorithm could be used to find lower-bounds for per-team SICs, which allows skipping the low-level search in many cases and can trivially be combined with pruned child generation branching within teams.

@@

	Another subject for future research might be a closer theoretical and experimental analysis of simple child pruning as alternative to enhanced pruning in the context of ICTS for MAPF, assuming this method has not been described before. Although experimentally it was shown to outperform enhanced pruning in exhaustive matching, ideally the two pruning strategies should be tested using the Moving AI benchmark \cite{sturtevant2012} to reach definite conclusions about how they compare. Using simple pruning to prune children and enhanced pruning for pruning low-level search could also be interesting to explore, as it could potentially combine the benefits of searching a smaller ICT and performing fewer $k$-agent low-level searches at the cost of spending more time pruning.
	\section{Responsible Research}
	The implementation of the ICTS-based algorithms described in this work is publicly available online in the form of a Github repository\footnote{\url{https://github.com/tbvanderwoude/icts-m}}. Included is also code for benchmarking different configurations of the ICTS solver and the maps that were used in the benchmarks reported above. The benchmarks and maps used in relative comparison of MAPFM algorithms is also available on Github\footnote{\url{https://github.com/jonay2000/research-project}}.
